---
title: "Data_7 Linear Models"
author: "Simon Willemin"
date: "(`r format(Sys.time(), '%d %B, %Y')`)"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(dslabs)
```

#The Dataset 

?cars

Let's get back to our faithful `cars` data. 


```{r str}
str(cars)
kable(head(cars))
```

This data - which is old (from the 1920's) - describes the speed of cars and the distances taken to stop. The dataframe has 50 observations of only two variables. Namely, the numeric speed (mph) and the stopping distance (ft). 

It is plain and intuitive to see here that speed influences the braking distance. Hence, speed will be our **explanatory variable*** and distance will be our **response variable**


*Scatterplot*

Let's take a visual lool at our data and plot the distance per speed : 

```{r  plot}
plot(cars, xlab = "Speed (mph)", ylab = "Stopping distance (ft)")
lines(lowess(cars$speed, cars$dist, f = 2/3, iter = 2), col = "blue")
title(main = "Stopping Distance (ft) / Speed (Mph)")
```


## Correlation?


Now we will take a look at if there is a correlation between speed and braking distance. 
In order to do that, we will conduct a Pearson's correlation. That is because both our predictor and output variables are continuous numerical variables. 

```{r pearson}
correlation <- cor(cars$dist, cars$speed, method = 'pearson')
print(correlation)
``` 

The correlation is strong, being *0.8068949*. 


## Visual representation 

```{r plot2}
library(ggplot2)
gg <- ggplot(cars, aes(speed,dist))
gg <- gg + geom_point()
gg <- gg + geom_smooth(alpha=0.3, method="lm")
print(gg)
```


We know that such positive correlation results do not mean causation, altough the relationship is quite obvious here. Let's hence run a general linear model. 

## General linear model 

### Assumptions 

**Independence of observations** -- > probably present 
**Normality** --> we will check it later on
**Linearity** --> we will see it in our linear model 
**Homoscedasticity** --> ""


###  Check fro normality of the distribution 
```{r normality}
hist((cars$dist), main="Cars distance",
     xlab ="distance")
hist((cars$speed), main="Cars speed",
      xlab = "speed")
```

It seems to be roughly normally around 20-40 



### Linearity
```{r linearity}
 plot(speed ~ dist, data = cars)
normality <- ggplot(cars, aes(speed,dist))
normality <- normality + geom_point()
normality <- normality + geom_smooth(alpha=0.3, method="lm")
print(normality)
```


## Simple regression
```{r regression}
speed.dist.lm <- lm(speed ~ dist, data = cars)
summary(speed.dist.lm)
```

This leads us to conclude that there is a significant positive relationship between speed and braking distance. 


# Reference
Ezekiel, M. (1930) Methods of Correlation Analysis. Wiley 

