---
title: " Data 4 - Data Visualization"
author: "Simon Willemin"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message= F , warning= F}
library(readr)
library(dslabs)
library(tidyverse)
library(knitr)
library(ggplot2)
```

We will be using this the `gapminder` dataset from the `dslabs` package which contains the following variables : 


* Country.
* Year.
* Infant_mortality --> Infant deaths per 1000.
* Life_expectancy -->  Life expectancy in years.
* Fertility --> Average number of children per woman.
* Population -->  Country population.
* GDP according to World Bank 
* Continent 
* Region -->  Geographical region.

What are the data types : 

```{r str}
str(gapminder)
```

--> 5 numeric vectors, 3 factors and 1 integer vector.

We will then transform it into a long format : 

```{r long}
gapminder_long <- gapminder %>% 
  pivot_longer(c("infant_mortality", "life_expectancy", "fertility", "population", "gdp"), names_to = "variable")
gapminder_long %>% head() %>% kable()
```

Then, we will proceed to unite both continent and region columns : 

```{r unite}
gapminder_long <- gapminder_long %>% 
  unite(cont_region, continent, region)
gapminder_long %>% head() %>% kable()
```

And then revert it : 

```{r separate}
gapminder_long <- gapminder_long %>% 
  separate(cont_region, into=c("continent", "region"), sep="_")
gapminder_long %>% head() %>% kable()
```

We will finally plot mortality against life expectancy : 

```{r plot, warning=FALSE}
ggplot(gapminder, aes(infant_mortality, life_expectancy)) +
  geom_point()
```


