
---
title: "Data Wrangling "
author: "Simon Willemin"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  github_document:
    toc: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(knitr)
library(dplyr)
library(readr)
library(dslabs)
```


## Filtering 

Let's first `filter()` the data for the United States and put them in a dataframe.
```{r}
USA_country_filter <- gapminder %>% 
  filter(country == "United States")
```

Let's now do that for Switzerland 

```{r}
Switzerland_country_filter <- gapminder %>% 
  filter(country == "Switzerland")
```

## Selecting

Let's now `select()` some interesting variable for the USA

```{r}
USA_selected <- USA_country_filter %>%
  select(fertility, population, year)
  
str(USA_selected)
head(USA_selected) %>% kable()
tail(USA_selected) %>% kable()
```

And the same for Switzerland 

```{r}
Switzerland_selected <- Switzerland_country_filter %>%
  select(fertility, population, year)
  
str(Switzerland_selected)
head(Switzerland_selected) %>% kable()
tail(Switzerland_selected) %>% kable()
```


## Arranging 
```{r}
USA_selected %>%
  arrange(year)
```

We can here notice that the overall fertility rates drop over the years, while the population constantly increases. 
Interestingly, after a lower-than-two fertility rate period in the 70's, it increased again later on, and stabilized.

```{r}
Switzerland_selected %>%
  arrange(year)
```

Whereas in Switzerland, it was at first (in the 1960's) lower than in the US at the same period. However, it shrank more regularly and is now at a low since the 1980's already ! 

## Group by and Summarize

```{r}
na.omit(gapminder) %>%
  group_by(country) %>%
  summarize(total_fertility = sum (fertility)) %>%
  arrange(desc(total_fertility)) %>%
  head()
```

```{r}
na.omit(gapminder) %>%
  group_by(country) %>%
  summarize(total_fertility = sum (fertility)) %>%
  arrange(desc(total_fertility)) %>%
  tail()
```

We can here usprisingly notice that the most fertile country (over the years) is an african country, namely Rwanda. Also, the least fertile one is Ireland. 

## Merging

We could for instance merge the `results_us_election_2016` and `murders` datasets. 

```{r}
head(full_join(results_us_election_2016, murders)) %>% kable()
```

# References

Jennifer Bryan (NA). gapminder: Data from Gapminder.
https://github.com/jennybc/gapminder,
http://www.gapminder.org/data/,
https://doi.org/10.5281/

Rogers, S. (2011, October 5). The murder map of London’s boroughs. The Guardian. http://www.theguardian.com/news/datablog/interactive/2011/oct/05/murder-london-map


WIlliams, Jack, 2022, "Elections Performance Index 2016", https://doi.org/10.7910/DVN/TSUQOA, Harvard Dataverse, V2, UNF:6:p1RbweyWRBJUm7A957ZaAQ== [fileUNF]


